---
title: "Reserve Study"
output:
  slidy_presentation:
    duration: 60
  ioslides_presentation: default
---


```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '~/GitHub/IntroR_CLRS2019')
```



# Project Overview

We will estimate reserves for an insurer's claims data set. At every step below, we will store the results in a list object which represents our reserve analysis of the insurer. We will also define functions to make our code reusable for another insurer. 

* Part 1 - Loading Claims Data  
* Part 2 - Data Aggregation 
* Part 3 - Helper Functions 
* Part 4 - Development Method  
* Part 5 - Initial Expected Loss Ratio method  
* Part 6 - Bornhuetter Ferguson Method  
* Part 7 - Your Turn
* Part 8 - Extra Credit!


# (1) Loading Claims Data

```{r }
claimsData <- read.csv("~/GitHub/IntroR_CLRS2019/slides_in/ppauto_pos_clean.csv",
                       stringsAsFactors = FALSE)
class(claimsData)
summary(claimsData)
```


```{r}
# GRCODE represents the code assigned to individual insurers by NAIC
claimsData$GRCODE <- as.factor(claimsData$GRCODE)
summary(claimsData$GRCODE)

# Let's analyze insurer with GRCODE 43
claimsData43 <- subset(claimsData, GRCODE == 43)
```

# (1) Loading Claims Data

## The origin trick

From the summary, you can see that the accident and transaction dates were read as integers. To convert it to date in R, we need to know the origin date which in Excel is December 30, 1899.

```{r }
# Convert columns to date
claimsData43$AccidentDate <- as.Date(claimsData43$AccidentDate,
                                     origin = "1899-12-30")
claimsData43$TransactionDate <- as.Date(claimsData43$TransactionDate,
                                        origin = "1899-12-30")
summary(claimsData43[, c("AccidentDate", "TransactionDate")])

# Calculate accident years and development ages
library(lubridate)
claimsData43$AccidentYear <- year(claimsData43$AccidentDate)
claimsData43$DevAge <- 12*(year(claimsData43$TransactionDate)-
                          year(claimsData43$AccidentDate)+1)
```

# (1) Loading Claims Data

Creating a function to compile data.

```{r}
# The function will take in 2 arguments
# data: claims data for all insurers
# insurerCode: the GR code of insurer that we want to analyze
# return: filtered dataset, with additional fields for triangle creation
compileData <- function(data, insurerCode){
  library(lubridate)
  # GRCODE represents the code assigned to individual insurers by NAIC
  data$GRCODE <- as.factor(data$GRCODE)

  # Let's analyze insurer with GRCODE = insurerCode
  insurerData <- subset(data, GRCODE == insurerCode)
  
  # Convert columns to date
  insurerData$AccidentDate <- as.Date(insurerData$AccidentDate,
                                      origin = "1899-12-30")
  insurerData$TransactionDate <- as.Date(insurerData$TransactionDate,
                                         origin = "1899-12-30")
  
  # Calculate accident years and development ages
  insurerData$AccidentYear <- year(insurerData$AccidentDate)
  insurerData$DevAge <- 12*(year(insurerData$TransactionDate)-
                            year(insurerData$AccidentDate)+1)
  
  # Output compiled data
  return(insurerData)
}
```

# (1) Loading Claims Data

Creating an insurer list object.

```{r}
# initializing the list object that represents insurer's reserve analysis
insurer43 <- list(data = NA)
# storing compiled data for insurer with GRCODE 43
insurer43$data <- compileData(data = claimsData, insurerCode = 43)
summary(insurer43$data)
```

# (2) Data Aggregation

Our loss triangles will be 10x10 matrices with 2 dimensions - accident year and development age (in months).

```{r}
# create paid triangle matrix
accYears <- sort(unique(insurer43$data$AccidentYear))
devAges <- sort(unique(insurer43$data$DevAge))
paidTri <- matrix(nrow = length(accYears), ncol = length(devAges),
                  dimnames = list(accYears, devAges))

# create cumulative paid loss triangle
for(i in 1:length(accYears)){
  for(j in 1:(length(devAges)-i+1)){
    paidTri[i,j] <- as.numeric(subset(insurer43$data,
                                     AccidentYear == accYears[i] & DevAge == devAges[j],
                                     CumPaidLoss))
    
  }
}
print(paidTri, na.print='')
```

# (2) Data Aggregation

Create a function to process triangles.

```{r}
# create function to create triangles
createTriangles <- function(data, triColName, triType){
  # create triangle matrix
  accYears <- sort(unique(data$AccidentYear))
  devAges <- sort(unique(data$DevAge))
  tri <- matrix(nrow = length(accYears), ncol = length(devAges),
                dimnames = list(accYears, devAges))
  
  # a matrix has different attributes
  # you can add a comment using the comment attribute
  # add comment on type of triangle
  comment(tri) <- triType
  
  # create cumulative paid loss triangle
  for(i in 1:length(accYears)){
    for(j in 1:(length(devAges)-i+1)){
      tri[i,j] <- as.numeric(subset(data,
                                    AccidentYear == accYears[i] & DevAge == devAges[j],
                                    triColName))
      
    }
  }
  
  return(tri)
}

# paid and reported loss triangle
insurer43$rptdTri <- createTriangles(insurer43$data, "IncurLoss", "Reported")
print(insurer43$rptdTri, na.print='')

insurer43$paidTri <- createTriangles(insurer43$data, "CumPaidLoss", "Paid")
print(insurer43$paidTri, na.print='')
```

# (2) Data Aggregation

Get and store earned premium data.

```{r}
# create function to get earned premium
getEarnPrem <- function(data){
  # get earned premium for accident years
  accYears <- sort(unique(data$AccidentYear))
  earnPrem <- sapply(accYears, 
                     function(x) unique(data$EarnedPremNet[data$AccidentYear == x]))
  # totals
  earnPrem <- c(earnPrem, sum(earnPrem))
  # add row names
  names(earnPrem) <- c(accYears, "Total")
  # output
  earnPrem  
}
insurer43$earnPrem <- getEarnPrem(insurer43$data)
insurer43$earnPrem
```

# (3) Helper Functions

Below, we will create some functions that will help us perform useful triangular manipulations.

```{r}
### function: lastDiag
### purpose: get last diagonal of a triangle, in vector form
### this is done by taking the tail non-NA value for each row
### arguments: tri = triangle matrix object
lastDiag <- function(tri) {
  apply(tri, MARGIN=1, FUN=function(x) tail(na.omit(x),1))
}


### function: ataTri
### purpose: get age-to-age triangle based on an input triangle
### arguments: tri = triangle matrix object
ataTri <- function(tri) {
  sapply(2:ncol(tri), function(x) tri[,x]/tri[,x-1])
}

### function: avgStr
### purpose: get unweighted A-A all year avg
### arguments: aaTri = age-to-age triangle matrix object
avgStr <- function(aaTri) {
  apply(aaTri, MARGIN=2, FUN=mean, na.rm=TRUE)
}

### function: avgWtd
### purpose: get weighted A-A all year avg
### arguments: tri = triangle matrix object
avgWtd <- function(tri) {
  aa <- ataTri(tri)
  tri_x_last <- tri[, 1:(ncol(tri)-1)]
  wtd <- aa * tri_x_last
  # mask for values to consider
  wtd_cell <- (!is.na(wtd))*1
  return(apply(wtd, MARGIN=2, FUN=sum, na.rm=TRUE) / 
           apply(wtd_cell * tri_x_last, MARGIN=2, FUN=sum, na.rm=TRUE))
}
```


# (4) Development Method

```{r}
# Chainladder development method
# Arguments:
# triangle = triangle matrix object to run the development method on
# paidTri = paid loss triangle matrix object to get unpaid loss
# tail (optional, default value = 1)
#         = tail factor to use in development method
# wtdAvg (optional, default value is not to use wtdAvg)
#         = boolean vector indicating use of weighted averages for each dev age
# Return:
#   list of results, including A-A vector, A-U vector, loss to date, 
devMethod <- function(triangle, paidTri, tail=1, wtdAvg=rep(FALSE, ncol(triangle))){
  # data frame to store dev method results
  results <- data.frame(accYear = c(row.names(triangle), "Total"),
                        latestLoss = NA, latestPaid = NA,
                        ata = NA, atu = NA, ult = NA, unpaid = NA)
  
  # get latest triangle losses and paid losses
  results$latestLoss <- c(lastDiag(triangle), NA)
  results$latestPaid <- c(lastDiag(paidTri), NA)
  
  # create ATA triangle
  ataTriangle <- ataTri(triangle)
  row.names(ataTriangle) <- row.names(triangle)
  colnames(ataTriangle) <- colnames(triangle)[-ncol(triangle)]
  
  # wtd all year avg
  ataWtd <- c(tail, rev(avgWtd(triangle)))
  # straight all year avg
  ataStr <- c(tail, rev(avgStr(ataTriangle)))
  # get selected ATA
  results$ata <- c(ifelse(wtdAvg, ataWtd, ataStr), NA)
  # get selected ATU
  results$atu <- cumprod(results$ata)
  
  # ultimate loss
  results$ult <- results$latestLoss*results$atu
  # unpaid loss
  results$unpaid <- results$ult - results$latestPaid
  
  # totals row - sums everything but last row (only for to-date and ultimate columns)
  results[nrow(results), c(2:3, 6:7)] <- sapply(c(2:3, 6:7),
                                                function(x) sum(results[-nrow(results), x]))
  
  # set column names
  colnames(results) <- c("Accident Year", paste("Latest", comment(triangle)),
                         "Latest Paid", "Age-to-Age", "Age-to-Ult",
                         "Ultimate Loss", "Unpaid Loss")
  
  # output results
  return(results)
}
```

# (4) Development Method

```{r}
# calculate and store dev method results for insurer 43
# paid dev method
insurer43$paidDev <- devMethod(triangle = insurer43$paidTri,
                               paidTri = insurer43$paidTri,
                               tail = 1)
# reported dev method
insurer43$rptdDev <- devMethod(triangle = insurer43$rptdTri,
                               paidTri = insurer43$paidTri,
                               tail = 1)

insurer43$paidDev
insurer43$rptdDev
```

# (5) Initial Expected Loss Ratio

```{r}
# IELR method
# Arguments:
# rptdDev = data frame object with reported development method results
# paidDev = data frame object with paid development method results
# earnPrem = earned premium
# rptdWt (optional, default value = 0.5)
#        = weight between 0 and 1 to use for reported ultimates 
#        = 1 implies full weight given to reported ultimates
#        = 0 implies full weight given to paid ultimates
#        = 0.5 implies equal weight given to paid and reported ultimates
# nyears (optional, default value = 5)
#        = number of recent years to include in IELR straight average
# excludeYears (optional, default value = 1)
#        = number of recent years to exclude in IELR straight average
IELR <- function(rptdDev, paidDev, earnPrem, rptdWt = 0.5, nyears=5, excludeYears=1){
  # data frame to store ielr method results
  results <- data.frame(accYear = names(earnPrem),
                        latestPaid = NA, earnPrem = earnPrem,
                        devLoss = NA, ultLR = NA, selLR = NA,
                        ult = NA, unpaid = NA)
  
  # get weighted ultimate losses for IELR
  results$devLoss <- rptdWt*rptdDev[,c("Ultimate Loss")] + 
                      (1-rptdWt)*paidDev[,c("Ultimate Loss")]
  
  
  # store latest loss, latest paid loss and earned premium
  results$latestPaid <- paidDev[, c("Latest Paid")]
  
  # calculate ultimate loss ratio
  results$ultLR <- results$devLoss / results$earnPrem
  
  # selected loss ratio as straight average of nyears LR
  ## calculate position of latest AY to include in IELR calculation
  latestAYPos <- nrow(results)-1-excludeYears
  results$selLR <- mean(results$ultLR[seq(latestAYPos-nyears+1, latestAYPos)])
  
  # ultimate and unpaid loss
  results$ult <- results$earnPrem * results$selLR
  results$unpaid <- results$ult - results$latestPaid
  
  # set column names
  colnames(results) <- c("Accident Year", "Latest Paid", "Earned Premium",
                         "Development Method Ultimate", "Ultimate Loss Ratio", "Selected Loss Ratio",
                         "IELR Ultimate Loss", "IELR Unpaid Loss")
  
  # output results
  return(results)
}
```

# (5) Initial Expected Loss Ratio

```{r}
# calculate and store IELR method results for insurer 43
insurer43$ielr <- IELR(rptdDev = insurer43$rptdDev, paidDev = insurer43$paidDev,
                       earnPrem = insurer43$earnPrem)
insurer43$ielr
```

# (6) Bornheutter-Ferguson Method

```{r}
# B-F Method
# Arguments:
# devResults = data frame object with development method results
# ielrResults = data frame object with IELR method results
# triType = "Paid" to use paid development patterns
#         = "Reported" to use reported development patterns
BF <- function(devResults, ielrResults, triType){
  # data frame to store ielr method results
  results <- data.frame(accYear = devResults$`Accident Year`,
                        latestLoss = NA, latestPaid = NA,
                        devPct = NA, devPctMod = NA, devEst = NA, ielrEst = NA,
                        ult = NA, unpaid = NA)
  
  # latest loss and latest paid loss
  results$latestLoss <- devResults[,paste("Latest", triType)]
  results$latestPaid <- devResults[,"Latest Paid"]
  
  # reported/paid %
  results$devPct <- 1/devResults[,"Age-to-Ult"]
  
  # B-F definition of credibility = reported/paid % is not valid
  # if reported/paid % > 1 i.e. when there is negative development
  results$devPctMod <- ifelse(results$devPct > 1, 1, results$devPct)
  
  # dev/ielr method estimates
  results$devEst <- devResults[, c("Ultimate Loss")]
  results$ielrEst <- ielrResults[, c("IELR Ultimate Loss")]
  
  # calculate B-F ultimate and unpaid loss
  results$ult <- results$devEst*results$devPctMod + results$ielrEst*(1-results$devPctMod)
  results$ult[nrow(results)] <- sum(results$ult[-nrow(results)])
  results$unpaid <- results$ult - results$latestPaid
  
  # set column names
  colnames(results) <- c("Accident Year", "Latest Loss", "Latest Paid",
                         paste(triType, "%"), paste(triType, "% for B-F"), "Development Method Ultimate",
                         "IELR Method Ultimate", "B-F Ultimate Loss",
                         "B-F Unpaid Loss")
  
  # output results
  return(results)
}
```

# (6) Bornheutter-Ferguson Method

```{r}
# calculate and store B-F method results for insurer 43
insurer43$paidBF <- BF(insurer43$paidDev, insurer43$ielr, "Paid")
insurer43$rptdBF <- BF(insurer43$rptdDev, insurer43$ielr, "Reported")
insurer43$paidBF
insurer43$rptdBF
```

# (7) Your turn

Use the functions we created to run a reserve analysis for another insurer. As a reference point, insurers are represented by the GRCODE variable in our claims data set. Following are the different GRCODEs you can use:

```{r}
unique(claimsData$GRCODE)
```

For your reference, following are the functions we created:

**_compileData(data, insurerCode)_**  
**_createTriangles(data, triColName, triType)_**  
**_getEarnPrem(data)_**  
**_lastDiag(tri)_**  
**_ataTri(tri)_**  
**_avgStr(aaTri)_**  
**_avgWtd(tri)_**  
**_devMethod(triangle, paidTri, tail, wtdAvg)_**  
    where _tail_ and _wtdAvg_ are optional arguments  
**_IELR(rptdDev, paidDev, earnPrem, rptdWt, nyears, excludeYears)_**  
    where _rptdWt, nyears, excludeYears_ are optional arguments  
**_BF(devResults, ielrResults, triType)_**  

# (8) Extra Credit!

If you are feeling overly ambitious today and want to improve your LDF selections, modify the devMethod() function we created to take in an argument nyearAvg, which would be a vector representing the number of recent years to include to get your LDF average for every development age.

For example, for a 10x10 triangle, nyearAvg would be a 1x9 numeric vector. If the first element of the vector is 5, the devMethod() function would calculate a 5-year average of 12-24 month ATAs.

Hint: In our function, you need to vary the start variable by the nyearAvg argument.
