---
title: "Reserve Study Solutions"
output:
  slidy_presentation:
    duration: 60
  ioslides_presentation: default
---


```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '~/GitHub/IntroR_CLRS2019')
```

```{r include=FALSE}
claimsData <- read.csv("~/GitHub/IntroR_CLRS2019/slides_in/ppauto_pos_clean.csv",
                       stringsAsFactors = FALSE)
source("actuarial_study_functions.R")
```

# (6) Your turn

Use the functions we created to run a reserve analysis for another insurer. As a reference point, insurers are represented by the GRCODE variable in our claims data set. Following are the different GRCODEs you can use:

```{r}
unique(claimsData$GRCODE)
```

For your reference, following are the functions we created:

**_compileData(data, insurerCode)_**  
**_createTriangles(data, triColName, triType)_**  
**_getEarnPrem(data)_**  
**_devMethod(triangle, paidTri, tail, wtdAvg)_**  
    where _tail_ and _wtdAvg_ are optional arguments  
**_IELR(rptdDev, paidDev, earnPrem, rptdWt, nyears, excludeYears)_**  
    where _rptdWt, nyears, excludeYears_ are optional arguments  
**_BF(devResults, ielrResults, triType)_**  

# (6) Your turn

We can use the functions we created to run a quick reserve analysis for another insurer.

```{r}
# initializing the list object that represents insurer's reserve analysis
insurer460 <- list(data = NA)
# storing compiled data for insurer with GRCODE 460
insurer460$data <- compileData(data = claimsData, insurerCode = 460)
summary(insurer460$data)
```

# (6) Your turn

```{r}
# create paid and reported loss triangle
insurer460$rptdTri <- createTriangles(insurer460$data, "IncurLoss", "Reported")
insurer460$paidTri <- createTriangles(insurer460$data, "CumPaidLoss", "Paid")
insurer460$rptdTri
insurer460$paidTri

# get earned premium for accident years
insurer460$earnPrem <- getEarnPrem(insurer460$data)
insurer460$earnPrem
```

# (6) Your turn

```{r}
# calculate and store dev method results for insurer 460
# paid dev method
insurer460$paidDev <- devMethod(triangle = insurer460$paidTri,
                               paidTri = insurer460$paidTri,
                               tail = 1)
# reported dev method
insurer460$rptdDev <- devMethod(triangle = insurer460$rptdTri,
                               paidTri = insurer460$paidTri,
                               tail = 1)

insurer460$paidDev
insurer460$rptdDev
```

# (6) Your turn

```{r}
# calculate and store IELR method results for insurer 460
insurer460$ielr <- IELR(rptdDev = insurer460$rptdDev, paidDev = insurer460$paidDev,
                       earnPrem = insurer460$earnPrem)
insurer460$ielr
```

# (6) Your turn

```{r}
# calculate and store B-F method results for insurer 460
insurer460$paidBF <- BF(insurer460$paidDev, insurer460$ielr, "Paid")
insurer460$rptdBF <- BF(insurer460$rptdDev, insurer460$ielr, "Reported")
insurer460$paidBF
insurer460$rptdBF
```

# (7) Extra Credit!

If you are feeling overly ambitious today and want to improve your LDF selections, modify the devMethod() function we created to take in an argument nyearAvg, which would be a vector representing the number of recent years to include to get your LDF average for every development age.

For example, for a 10x10 triangle, nyearAvg would be a 1x9 numeric vector. If the first element of the vector is 5, the devMethod() function would calculate a 5-year average of 12-24 month ATAs.

Hint: In our function, you need to vary the start variable by the nyearAvg argument.

# (7) Extra Credit!

```{r}
# Chainladder development method
# Arguments:
# triangle = triangle matrix object to run the development method on
# paidTri = paid loss triangle matrix object to get unpaid loss
# tail (optional, default value = 1)
#         = tail factor to use in development method
# wtdAvg (optional, default value is not to use straight average)
#         = boolean vector indicating use of weighted averages for each dev age
# nyearAvg (optional, default value is not to use all years)
#         = numeric vector giving the number of recent years to include
#           to get your LDF average for every development age
devMethod <- function(triangle, paidTri, tail=1, wtdAvg=rep(FALSE, ncol(triangle)-1),
                      nyearAvg=rep(nrow(triangle)-1, ncol(triangle)-1)){
  # data frame to store dev method results
  results <- data.frame(accYear = c(row.names(triangle), "Total"),
                        latestLoss = NA, latestPaid = NA,
                        ata = NA, atu = NA, ult = NA, unpaid = NA)
  
  # create ATA triangle
  ataTri <- sapply(2:ncol(triangle),
                   function(x) triangle[,x]/triangle[,x-1])
  row.names(ataTri) <- row.names(triangle)
  colnames(ataTri) <- colnames(triangle)[-ncol(triangle)]
  
  # ATA/ATU for oldest year will be our tail factor
  results$ata[1] <- tail
  results$atu[1] <- tail
  
  # latest loss for oldest year
  results$latestLoss[1] <- triangle[1, ncol(triangle)]
  # latest paid loss for oldest year
  results$latestPaid[1] <- paidTri[1, ncol(paidTri)]
  
  # calculate ATA/ATU for remaining years
  for(j in seq(ncol(triangle), 2, -1)){
    # calculate position of latest AY for dev age j
    i <- nrow(triangle) - j + 1 
    # EDIT: calculate starting and ending index for number of years to take average of
    end <- i
    if(nyearAvg[j-1] > end){
      start <- 1
    } else{
      start <- end - nyearAvg[j-1] + 1
    }
    # ATA for AY i+1 from age j-1 to j (all year weighted average)
    if(wtdAvg[j-1]){
      # all year weighted average ATA
      results$ata[i+1] <- sum(ataTri[start:end,j-1]*triangle[start:end,j-1]) /
                          sum(triangle[start:end,j-1])
    } else{
      # all year straight average ATA
      results$ata[i+1] <- mean(ataTri[start:end,j-1])
    }
    
    # ATU for AY i+1 from age j-1 to ult
    results$atu[i+1] <- results$atu[i]*results$ata[i+1]
    # latest loss for AY i+1
    results$latestLoss[i+1] <- triangle[i+1, j-1]
    # latest paid loss for AY i+1
    results$latestPaid[i+1] <- paidTri[i+1, j-1]
  }
  
  # ultimate loss
  results$ult <- results$latestLoss*results$atu
  # unpaid loss
  results$unpaid <- results$ult - results$latestPaid
  
  # totals
  results[nrow(results), c(2:3, 6:7)] <- sapply(c(2:3, 6:7),
                                                function(x) sum(results[-nrow(results), x]))
  
  # set column names
  colnames(results) <- c("Accident Year", paste("Latest", comment(triangle)),
                         "Latest Paid", "Age-to-Age", "Age-to-Ult",
                         "Ultimate Loss", "Unpaid Loss")
  
  # output results
  return(results)
}
```

# (7) Extra Credit!

```{r}
# calculate and store dev method results for insurer 460
# paid dev method
insurer460$paidDev <- devMethod(triangle = insurer460$paidTri,
                                paidTri = insurer460$paidTri,
                                tail = 1,
                                nyearAvg = rep(5, ncol(insurer460$paidTri)-1))

insurer460$paidDev
```