---
title: Functions
author: "Steve Berman"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ./css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: no
    theme: solarized
    transition: slide
---

```{r echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
    warning=TRUE
  , error=TRUE
  , echo=TRUE
  , message=FALSE
  , fig.height = 4.5
  , fig.pos="t"
  , collapse = TRUE
)
knitr::opts_knit$set(root.dir = normalizePath('../'))
```

# Functions, part 1

## What's the big deal?

```{r eval=FALSE}
some_var <- sqrt(pi) + 5
some_function <- function(x) {
  sqrt(x) + 5
}
some_function(pi)
```

## Here's what's amazing

1. Functions expand your toolset
2. Functions are __objects__, just like everything else in R
3. Functions can return any value
4. 1 and 2 mean that functions can return other functions
5. Many functions are __vectorized__ (that is, they take vectors and return vectors)
6. Functions may be altered after they're created

# Basic functions

## R has a great variety of functions that

* Perform mathematical calculations
* Convert from one data type to another
* Provide formatting
* Load data, save data, manipulate data
* Many more!

# What is a function made of?

## A simple user-defined function

```{r}
layer <- function(loss, attachment, limit, share){
  share * pmin(limit, pmax(0, loss - attachment))
}
```

## The elements of a function

1. arguments
2. body
3. environment - the scope that any object exists in

# Parameters

## Parameters

```{r}
# formals function allows access and manipulation of formal arguments
formals(layer)
```

Note that it _doesn't_ tell you what data types they are. R is not strongly typed.

## Matching arguments

1. Exact matching - actuals and formals must be 1 to 1
2. Partial matching - (on unmatched) actuals matching the first part of the formals - cannot have multiple matches
3. Positional matching - unmatched paired with unnamed

## Matching arguments

```{r}
myFunc <- function(mojo, monkey, banana){
  print(match.call())   # match.call returns a call in which all arguments are specified by their full names
}

myFunc(mojo = 1, mo = 2, 3)  # exact, then partial, then positional
myFunc(mo = 2, mojo = 1, 3)  # order not important - exact matching occurs first
myFunc(mo = 1, mo = 2, 3)    # CRASH on partial match
myFunc(1, 2, 3)              # all positional
```

# Functions, part 2

## Optional values (with defaults)

Allow you to have simple calls or provide a lot more detail

```{r}
myFunc <- function(a, b = 15) {
  a + b
}
myFunc(3)
```

No value supplied for the `b` argument, so default used.

## Check inputs

You can use the `missing()` function for this.

```{r}
myFunc <- function(mojo, monkey, banana){
  if (missing(banana)) {
    warning("I can't think without bananas")
    banana = monkey + 5
  }
}

myFunc(1, 2, 3)
myFunc(1, 2)
```

## Aside

* `stop()`: This is an error. The program can't go on.
* `warning()` - This tells the user that something happened and they might not like it, but it's not a critical issue.
* `message()` - Just sayin' something you might want to hear.


## Restrict inputs to set choice

Where a parameter may have only a small logical set of values.

```{r}
myFunc <- function(x, dist = c("lognormal", "gamma")){
  # argument verification using partial matching
  dist = match.arg(dist)
  print(dist)
}
myFunc(1:10, "lo")
myFunc(1:10, "jo")

myOtherFunc <- function(x, dist = c("lognormal", "gamma")){
  print(dist)
}
myOtherFunc(1:10, "lo")
myOtherFunc(1:10, "jo")
myOtherFunc(1:10, c("lo", "jo"))
```

## Defaults and missing arguments - best practices

* If you _must_ know what the user intends, _or_ if the default requires some logic, don't add a default.
* If you're fine with common default parameters (i.e. the base of the log function is e or 10), add a default.
* Don't add a default and check for a missing value. It's confusing and unnecessary.

## Ellipses

`...` is used to pass unspecified arguments to other functions. Typically used if your function sits between the user and some other function like `plot`

```{r, echo=TRUE}
dfData <- data.frame(PolicyYear = 2001:2010, Premium = rnorm(10))
myPlotFunc <- function(dfData, ...){
  plot(dfData$PolicyYear, dfData$Premium, ...)
}
```

```{r eval=FALSE}
myPlotFunc(dfData, pch=18)
myPlotFunc(dfData, pch=23, col="red")
```

```{r}
formals(myPlotFunc)
```

# Return value

## What does a function return?

Anything.

## return()

* `return()` isn't required. Good practice, though.
* A function will always return the last evaluated expression.
* `return()` is often used for _abnormal_ returns

It's good practice to use return statement as this is explicit

```{r}
myFunc <- function(x) {
  if (is.na(x)) return("Value missing")
  
  sum(x)
}
```


## Exercise

* $\mu=(E[X])- 0.5 * sigma^2$
* $\sigma^2=ln(1+cv^2)$

Create a function called LognormalParams, which will translate the mean and CV into parameters for the lognormal distribution. Test it using the following statements:

```{r eval=FALSE}
LognormalParams(10e3, 3.0)
LognormalParams(c(10e3, 20e3), 3.0)
LognormalParams(10e3)
LognormalParams(CV = 4, 10e3)
```

## Solution

```{r}
LognormalParams <- function(mean, CV) {
  sig2 <- log(1+CV**2)
  sig <- sqrt(sig2)
  mu <- mean - 0.5 * sig2
  return(c(mu, sig))
}
```

# If we have time

## Replacement functions

These take the form "xxx<-".  In the example below, the first argument is modified in place.

(These examples are virtually identical to the ones in Hadley's "Advanced R")

```{r}
`mojo<-` <- function(x, value){
  x <- x + value
} 
gonzo <- 1:10
gonzo
mojo(gonzo) <- 5
gonzo
```

Why?  In some cases, the format is easier to understand:
```{r eval=FALSE}
names(x) <- c("field1", "field2", "field3")
```

## Replacement functions (2)

```{r}
`mojo<-` <- function(x, position, value){
  x[position] <- x[position] + value
  x
} 
gonzo <- rep(1, 10)
gonzo
mojo(gonzo, 2) <- 5
gonzo
mojo(gonzo, 2:3) <- 5
gonzo
```

## Infix functions

Typically R functions are postfix (that is, the parameters come after the function definition)

```{r}
`%l^%` <- function(a, b){
  log(sqrt(a +b))
}
5 %l^% 3
log(sqrt(5+3))
```


