---
title: Getting Started
output:
  revealjs::revealjs_presentation:
    center: no
    css: ../css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: false
    theme: solarized
    transition: slide
---

```{r include=FALSE}
source('common.R')
```

##  Options

### Available in Tools/Global Options

* General: Save of history, or .Rdata (probably best if you don't)
* General: Default version of R
* Appearance: Choose editor themes - can create custom
* Code / Diagnostics
* Pane Layout: Change panel positions
* Code / Completion: snippets, autocompletion
    
## Script

* Create a script
* Save it, close it, open it

## Keyboard shortcuts

https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts

Keyboard Shortcut | Function
------------------| -----------------------------------------
ALT+SHIFT+K       | Show quick reference
ALT+-             | Insert assignment operator (`<-`)
CTRL+ENTER        | Execute selected code
CTRL+1            | Go to Source Editor
CTRL+2            | Go to Console
CTRL+SHIFT+M      | Insert pipe operator (`%>%`)
CTRL+L            | Clear console
CTRL+SHIFT+N      | Create new file
CTRL+W            | Close active document
CTRL+SHIFT+F      | Find in files
CTRL+SHIFT+B      | Build / rebuild package

## Code execution

CTRL+ENTER

90% of your code writing and execution will be done in scripts. Use the console for getting help or one-off operations like: package installation, light debugging.

## Speaking of debugging

R has a debugging framework....



## The working directory

```{r}
getwd()
```

You will probably never have a good reason to call `setwd()`. Use a project instead.

## Projects (*.Rproj)

* Collect all of the files relevant to a project
* Automatically set the working directory 
* Intermediate stuff:
  * Version control (with Git)
  * Package development

## Packages

There are thousands!

```{r eval=FALSE}
install.packages('actuar')
library(actuar)
```

Can install through Tools / Install packages as well.

## Data in packages

```{r eval=FALSE}
data(package = 'Lahman')   # load or list available datasets
data(iris)
```

datasets packages contains some sample data

## Installation vs. library

* install.packages()` copies files to your computer.
* `library()` finds those files and loads them into memory, where they remain until the session ends

## Package installation folder

```{r}
.libPaths()
```

Create a "universal" library by creating a file called ".Renviron".

```
R_LIBS="~/Rlibs"
```

https://stat.ethz.ch/R-manual/R-devel/library/base/html/Startup.html

## Exercise

Create an RStudio project to contain your work for this training.
