---
title: "R basics"
output:
  slidy_presentation:
    duration: 45
    fig_height: 3
    fig_width: 6
  beamer_presentation:
    fonttheme: professionalfonts
---

# Basics

## Assignment

Use the <code class="remark-inline-code"><-</code> operator.  The equal sign also works.

## Libraries
* <code class="remark-inline-code">library()</code>: see a list of available packages
* Use that to load a library as well: <code class="remark-inline-code">library(MASS)</code>

## Console window

* Can run individual commands, including retrieving or changing existing variables.
* Good for debugging.
* Longer programming is done through scripts.

## Get help!

* <code class="remark-inline-code">?</code> brings up help on a specific function
* <code class="remark-inline-code">??</code> performs a search on a term, if you don't know the exact name

## Mathematical operations:

* Addition, subtraction, etc. all work as expected
* Use `**` for exponentiation, `**` for modulus, `//` for integer division, etc. 

## Functions:
* Called similar to Excel:
<code class="remark-inline-code">abs(-3.5)</code> --> 3.5
* Parentheses required even if no parameters
* Some functions have optional parameters - can enter parameters in order or reference by name

## Comments, whitespace:
* Comments start with `#` (no multiple line comment)
* White space, indentations ignored by R

## Workspaces:
* Use Load Workspace / Save Workspace for saving and restoring objects (variables and functions) but not scripts

# Other starting considerations:

* Primitives (<code class="remark-inline-code">numeric, logical, character</code>)
* Other data types (<code class="remark-inline-code">list, vector, matrix, factor, "data.frame"</code>)
* The structure function (<code class="remark-inline-code">str</code>): show structure of an R object</li>
* The <code class="remark-inline-code">tidyverse</code>: a small universe of really useful packages (<code class="remark-inline-code">install.packages(pkgs='tidyverse')</code>)</li>
  <li>Styleguide <a href="http://adv-r.had.co.nz/Style.html">http://adv-r.had.co.nz/Style.html</a></li>
</ol>

# Read Data

<ol>
  <li> Comma-separated values (<code class="remark-inline-code">read.csv, readr::read_csv</code>)</li>
  <li> Comma-separated values (<code class="remark-inline-code">readxl::read_excel</code>)</li>
</ol>

Other data types are readable.  There are packages for SAS, SQL, Word, PDF, etc.

# Exercise 1

Locate the <code class="remark-inline-code">mtcars.csv</code> file that comes with <code class="remark-inline-code">readr</code>
```{r}
library(readr)
readr_example("mtcars.csv")
```
<exercise>Read this file using <code class="remark-inline-code">read.csv</code> and assign this to object <code class="remark-inline-code">cas01</code><br>
<exercise>Read this file using <code class="remark-inline-code">read_csv</code> and assign this to object <code class="remark-inline-code">cas02</code><br>
What is the difference between the files?
</exercise>

# Solution 1

```{r, message=FALSE}
fn <- readr_example("mtcars.csv")
cas01 <- read.csv(file = fn)
cas02 <- read_csv(file = fn)
class(cas01)
class(cas02)
```

# Exercise 2

Locate the <code class="remark-inline-code">mtcars.csv</code> file that comes with <code class="remark-inline-code">readr</code>
```{r}
library(readxl)
readxl_example('deaths.xlsx')
```
<exercise>Read this file using <code class="remark-inline-code">read_excel</code> and assign this to object <code class="remark-inline-code">cas03</code><br>
How many row and columns does this dataframe contain?
</exercise>

# Solution 2: Try #1

```{r, message=FALSE}
fn <- readxl_example('deaths.xlsx')
cas03 <- read.csv(fn)
str(cas03)
```

# Solution 2: Try #2

```{r}
fn <- readxl_example('deaths.xlsx')
cas03 <- read_excel(fn, sheet = 'arts' , range = 'a5:f15')
str(cas03)
```


# Read in CAS Data
```{r}
#auto_dat_cas <- read_csv('http://www.casact.org/research/reserve_data/ppauto_pos.csv')
```




