---
title: "Visualization"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ../css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: false
    theme: solarized
    transition: slide
---

```{r include=FALSE}
source('common.R')
```

# Base plotting

## Base plotting (1)

```{r}
library(raw)    # R Actuarial Workshop
data("RegionExperience")
RegionExperience$Region_f <- factor(RegionExperience$Region)

head(RegionExperience, 3)

```

## Base plotting (2)

```{r}
# basic plotting
plot(RegionExperience$PolicyYear, RegionExperience$NumClaims)
```


## Add labels

```{r}
# some more customization
# ex: pch affects how the "dot" looks
plot(RegionExperience$PolicyYear
  , RegionExperience$NumClaims
  , pch = 19
  , xlab = 'Policy Year'
  , ylab = '# of claims')
```

## Some more customization

```{r}
# some more customization
# ex: pch affects how the "dot" looks
plot(RegionExperience$PolicyYear
  , RegionExperience$NumClaims
  , col = RegionExperience$Region_f
  , pch = 19
  , xlab = 'Policy Year', ylab = '# of claims'
  , main = 'Claim counts', ylim=c(0, 10000))
```


## Base plotting

* Kinda ugly
* Customization is possible, but tedious

# ggplot

## ggplot2

gg =  "grammar of graphics" based on Leland Wilkinson

1. Data
2. Mapping
2. Geometric layers
3. Scales

# ggplot - Data

## ggplot - Data

```{r error = TRUE}
library(ggplot2)
library(dplyr)
plt_base <- ggplot(RegionExperience)

plt_base

data(COTOR2)
ggplot(COTOR2)
```

## ggplot - Data

Data must use a data frame.

```{r}
ggplot(1:10)
```

# ggplot - Mapping

## ggplot - Mapping

Mappings bind data to visual elements. Every data-derived visual element must be mapped to data.

```{r }
# aes = aesthetic mapping
plt_base <- plt_base + aes(x=PolicyYear, y=NumClaims)

plt_base <- ggplot(RegionExperience, aes(PolicyYear, NumClaims))
```

# ggplot - Layers

## Adding layers

Although we have data and we've mapped to elements of a coordinate system, we haven't specified what the visual elements should be. The `geom_*` family of functions add geometric shapes. 

```{r }
plt_base <- plt_base + geom_point()
plt_base
```

## One step

```{r }
# note pipe operator
RegionExperience %>% 
  ggplot(aes(x=PolicyYear, y=NumPolicies)) + 
  geom_point()
```

## Try multiple layers

```{r }
RegionExperience %>% 
  ggplot(aes(x=PolicyYear, y=NumPolicies)) + 
  geom_point() + 
  geom_line()
```

## Try that again

```{r }
RegionExperience %>% 
  ggplot(aes(x=PolicyYear, y=NumPolicies)) + 
  geom_point() + 
  geom_line(aes(color=Region))
```

## Statistical smoothers

```{r}
RegionExperience %>% 
  ggplot(aes(x=PolicyYear, y=NumPolicies)) + 
  geom_point() + 
  geom_smooth(method=lm, aes(color=Region))
```

# Scales

## Store a basic plot in an object

```{r}
plt <- RegionExperience %>% 
  ggplot(aes(x=PolicyYear, y=NumClaims)) + geom_point()
```

## Change the y scale

```{r}
plt + scale_y_continuous(labels = scales::comma)
```

## Change the x breaks

```{r}
plt + scale_x_continuous(breaks = 2001:2010)
```

## Change the x breaks

```{r}
my_breaks <- function(lims) {
  seq(ceiling(lims[1]), floor(lims[2]), by=1)
}
plt + scale_x_continuous(breaks=my_breaks)
```

# Facets

## Facets

Facets split the data into groups and draws a different plot for each group.

```{r}
ggplot(RegionExperience, aes(x=PolicyYear, y=NumClaims, color=Region)) + 
  geom_point() + facet_wrap(~ Region)
```

# Non-data visual elements

## Non-data visual elements

* labels
* gridlines
* legend
* axis ticks

## Labels

```{r}
plt + labs(x = 'Policy year', y = '# of claims')
```

## Gridlines, legends, axis ticks

TODO: complete this section

```{r}

```



## Stuff you can do with ggplot:

Line chart, bar chart, boxplot, scatterplots, dumbbell plot, violin plot, heatmaps, trellis plots, slopecharts, more!

## Exercises

1. Create a scatter plot for policy year and number of claims
2. Color each point based on region
3. Add a linear smoother. Which region is showing the greatest increase in claims?
4. Form the policy frequency by taking the ratio of claims to policies. Plot this.

Extra credit:

Use the state data to create a time series number of claims. Facet by region.

## Answers:

```{r}
library(raw)
data("RegionExperience")
plt1 <- ggplot(RegionExperience, aes(x = PolicyYear, y = NumClaims)) + geom_point()
plt1

plt2 <- plt1 + aes(color=Region)
plt2

plt3 <- plt2 + stat_smooth(method = "lm")
plt3

RegionExperience$Frequency <- with(RegionExperience, NumClaims / NumPolicies)

plt4 <- ggplot(RegionExperience, aes(x=PolicyYear, y=Frequency, color=Region)) + geom_point()+ 
  geom_line() + stat_smooth(method=lm)
plt4
```

```{r} -->
data("StateExperience")
pltExtra <- ggplot(StateExperience, aes(x=PolicyYear, y=NumClaims, color=Postal)) +
   geom_point() + geom_line()
pltExtra + facet_wrap(~ Region)
```

## Summary

* `ggplot2` is different at first, but will repay your investment.
* Works very well with grouped data to color/facet points.
* Fine-tuning things like axis labels can be a headache, but will get easier.
